---
title: "Market Basket Analysis"
author: "Andreu Oros, Sergi Pallice, Joël Ribera"
date: "`r Sys.Date()`"
output:
  pdf_document:
    df_print: kable
    toc: true
    number_sections: false
---


```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="150")
opts_chunk$set(echo=T,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
path <- getwd()

knitr::opts_knit$set(root.dir = normalizePath(path.expand(path),
                                             winslash = ("/"),
                                             mustWork = TRUE))
opts_knit$set(width=75)
```

## Executive Summary
### Introduction
Danielle has asked the team to perform a market basket analysis to help Blackwell's board of directors better understand the clientele that Electronidex is currently serving and if Electronidex would be an optimal acquisition.

A dataset of transactions has been provided. The dataset contains 9835 transactions and 125 different products over a 30-day period, or about 327 transactions a day. This tells us the retailer is neither large, nor small.


### Results and conclussions
After the analysis, we conclude Electronidex's sales are to be categorized in two forms: retail (B2C) and corporate (B2B). Had we had this information previous to our analysis, it would have saved time in the exploration phase.

Interesting patterns and item relationships found:
Retail:
>



Would Blackwell benefit in selling Electronidex's items?




### Limitations and observations:
 Properties of the dataset:
 1. The iMac is the product most bought, in 20% of all transactions. This high number stands out considering the large variety of products, especially being the iMac a pricey product. If this number is representative of all sales throughout the year, then Electronidex is potentially profitable.
 2. The mean of items bought per transaction is almost 5. Logically, I would say most people in the real world would buy 1 or 2 items per transactions most frequently, in an eletronics store.


### Recommendations
**Blackwell should acquire Electronidex**
If Blackwell does acquire Electronidex, do you have any recommendations for Blackwell? (Ex: cross-selling #items, sale promotions, should they remove items, etc.)


## Loading Packages and importing datasets

```{r, warning = FALSE, message = FALSE, results='hide'}
pacman::p_load(readr,rstudioapi,ggplot2,party,dplyr,arules,arulesViz,
               RColorBrewer,readxl,tidyr,Hmisc)
setwd("..")
trans <- read.transactions("./Datasets/ElectronidexTransactions2017.csv", 
                  format = "basket",sep=",", rm.duplicates=TRUE)
labels <- read_excel("./Datasets/ElectronidexItems2017.xlsx")
```

## Preprocessing

```{r}
labels[is.na(labels)] <-"Unknown"
summary(trans)
trans@itemInfo$labels<- labels$ProductType
trans@itemInfo$category <- labels$Category

trans_matrix <- as(trans,"matrix")
trans_df <- as.data.frame(trans_matrix)
#Turning into a sparematrix of 1s and 0s.
for (i in 1:ncol(trans_df)){
  trans_df[,i] <- as.integer(trans_df[,i])
}

```

## Feature Engineering

```{r}
#Counting number of items
nitems <- c()
for (i in 1:nrow(trans_df)) {
  nitems <- c(nitems, sum(trans_df[i,]))
}
trans_df$nitems <- nitems
trans_df$laptops <- trans_df[,which(colnames(trans_df) == "LG Touchscreen Laptop")] +
                      trans_df[,which(colnames(trans_df) == "Acer Aspire")] +
                      trans_df[,which(colnames(trans_df) == "HP Laptop")] +
  trans_df[,which(colnames(trans_df) == "ASUS Chromebook")] +
  trans_df[,which(colnames(trans_df) == "Apple Macbook Pro")] +
  trans_df[,which(colnames(trans_df) == "Apple MacBook Air")] +
  trans_df[,which(colnames(trans_df) == "Dell Laptop")] +
  trans_df[,which(colnames(trans_df) == "Eluktronics Pro Gaming Laptop")] +
  trans_df[,which(colnames(trans_df) == "Alienware AW17R4-7345SLV-PUS 17\" Laptop")] +
  trans_df[,which(colnames(trans_df) == "HP Notebook Touchscreen Laptop PC")]
trans_df$desktop <- trans_df[,which(colnames(trans_df) == "Lenovo Desktop Computer")] +
  trans_df[,which(colnames(trans_df) == "iMac")] +
  trans_df[,which(colnames(trans_df) == "HP Desktop")] +
  trans_df[,which(colnames(trans_df) == "ASUS Desktop")] +
  trans_df[,which(colnames(trans_df) == "Dell Desktop")] +
  trans_df[,which(colnames(trans_df) == "Intel Desktop")] +
  trans_df[,which(colnames(trans_df) == "Acer Desktop")] +
  trans_df[,which(colnames(trans_df) == "CYBERPOWER Gamer Desktop")] +
  trans_df[,which(colnames(trans_df) == "Dell 2 Desktop")] 
trans_df$tablet <- trans_df[,which(colnames(trans_df) == "iPad")] +
  trans_df[,which(colnames(trans_df) == "iPad Pro")] +
  trans_df[,which(colnames(trans_df) == "Fire HD Tablet")] +
  trans_df[,which(colnames(trans_df) == "Samsung Galaxy Tab")] +
  trans_df[,which(colnames(trans_df) == "Kindle")]
trans_df$printer <- trans_df$`Epson Printer`+ trans_df$`HP Wireless Printer` +
  trans_df$`Canon Office Printer` + trans_df$`Brother Printer` + 
  trans_df$`DYMO Label Manker`
trans_df$nmain <- trans_df$printer + trans_df$laptops + trans_df$desktop +
  trans_df$tablet
trans_df$ncomp <- trans_df$nitems - trans_df$nmain
trans_df$value <- 10*trans_df$nmain + trans_df$ncomp
```

## Visualizations 

```{r, warning = FALSE, message= FALSE}
products <- c("laptops", "desktop","printer","tablet")
for (i in products){
  print(ggplot(trans_df, aes_string(x = i)) + 
          geom_bar(fill = "lightgreen", bins = 100) + 
           
      ggtitle(paste("Histogram of",i)))
}
```

```{r}
for (i in products){
  for (j in products[-1]){
    if (i != j | j !=i) {
       print(ggplot(trans_df, aes_string(x = i, y = j, color = i)) +    geom_jitter()
          +
      ggtitle(paste("Scatterplot of",j,"vs",i))) 
      
    }
    
  }
}
```

```{r}
ggplot(trans_df, aes(x = nmain, y = ncomp)) +
  geom_jitter(color = "lightgreen")

```

## Splitting dataframe between corporates and retailers

```{r}
#Filtering. Corporate will be those transactions with more than 2 main products and 3 complements.
corporate <- filter(trans_df, nmain >= 2 | ncomp >= 3)
#Cleaning the new columns we created before
corporate <- corporate[,-which(colnames(corporate) %in% c("laptops","desktop","printer","tablet","nitems","nmain","value","ncomp"))]
#Filtering. Retailers will be those transactions with less than 2 main products and 3 complements.
retailer <- filter(trans_df,nmain <= 1 & ncomp <= 2)
#Cleaning
retailer <- retailer[,-which(colnames(retailer) %in% c("laptops","desktop",
"printer","tablet","nitems","nmain","value","ncomp"))]
#Transforming the dataframe into a transaction objecte
trans_corp <- as(corporate == 1, "transactions")
trans_retail <- as(retailer == 1, "transactions")
#Inserting labels and the level category
trans_corp@itemInfo$labels <- labels$ProductType
trans_corp@itemInfo$category <- labels$Category
trans_retail@itemInfo$labels <- labels$ProductType
trans_retail@itemInfo$category <- labels$Category
```

```{r}
itemFrequencyPlot(trans_corp,topN=10,type="relative",col=brewer.pal(8,'Pastel2'), 
                  main="Corp Relative Item Frequency Plot")

itemFrequencyPlot(trans_retail,topN=10,type="relative",col=brewer.pal(8,'Pastel2'), 
                  main="Retail Relative Item Frequency Plot")
```

## Creating rules via apriori algorithm

### Rules for products in corporate transactions

```{r, warning=FALSE,message=FALSE,results='hide'}
rules_corpro <- apriori (trans_corp, parameter = list(supp = 0.01, conf = 0.01, 
                                                 minlen = 2, maxlen = 4))
rules_corpro <- rules_corpro[!is.redundant(rules_corpro)]
```


```{r, warning=FALSE,message=FALSE}
summary(rules_corpro)
#The most populars products
corpro_sup <- inspect(head(sort(rules_corpro, by ="support"),10))
#Items that have high chances of being bought together
corpro_con <- inspect(head(sort(rules_corpro, by ="confidence"),10))
#Lift
corpro_lift <- inspect(head(sort(rules_corpro, by ="lift"),10))
```

### Rules for categories in corporate transactions

```{r, warning=FALSE,message=FALSE,results='hide'}
trans_corcat <- aggregate(trans_corp, by = "category")
rules_corcat <- apriori(trans_corcat, parameter = list(supp = 0.01, conf = 0.01, 
                                                 minlen = 2, maxlen = 4))
rules_corcat <- rules_corcat[!is.redundant(rules_corcat)]
```


```{r, warning=FALSE,message=FALSE}
summary(rules_corcat)
#The most populars categories
corcat_sup <- inspect(head(sort(rules_corcat, by ="support"),10))
#Items that have high chances of being bought together
corcat_con <- inspect(head(sort(rules_corcat, by ="confidence"),10))
#Lift
corcat_lift <- inspect(head(sort(rules_corcat, by ="lift"),10))
```
### Rules for products in retailers transactions

```{r, warning=FALSE,message=FALSE,results='hide'}
rules_retpro <- apriori(trans_retail, parameter = list(supp = 0.001, conf = 0.01, 
                                                 minlen = 2, maxlen = 4))
rules_retpro <- rules_retpro[!is.redundant(rules_retpro)]
```


```{r, warning=FALSE,message=FALSE}
summary(rules_retpro)
#The most populars products
retpro_sup <- inspect(head(sort(rules_retpro, by ="support"),10))
#Items that have high chances of being bought together
retpro_con <- inspect(head(sort(rules_retpro, by ="confidence"),10))
#Lift
retpro_lift <- inspect(head(sort(rules_retpro, by ="lift"),10))
```

### Rules for categories in retailers transactions

```{r, warning=FALSE,message=FALSE,results='hide'}
trans_retcat <- aggregate(trans_retail, by = "category")
rules_retcat <- apriori(trans_retcat,parameter = list(supp = 0.01, conf = 0.01, 
                                                      minlen = 2, maxlen = 4))
rules_retcat <- rules_retcat[!is.redundant(rules_retcat)]
```

```{r, warning=FALSE,message=FALSE}
summary(rules_retcat)
#The most populars products
retcat_sup <- inspect(head(sort(rules_retcat, by ="support"),10))
#Items that have high chances of being bought together
retcat_con <- inspect(head(sort(rules_retcat, by ="confidence"),10))
#Lift
retcat_lift <- inspect(head(sort(rules_retcat, by ="lift"),10))
```

## Rules visualizations

```{r, warning=FALSE, message=FALSE}
methods <- c("graph","scatterplot")
for (i in methods) {
  plot(rules_corpro, method = i,control=list(type="items"), max = 10, 
       main = paste(capitalize(i),"of rules of products bought by corporates"))
  plot(rules_corcat, method = i,control=list(type="items"), max = 10,
       main = paste(capitalize(i),"of rules of categories bought by corporates"))
  plot(rules_retpro, method = i,control=list(type="items"), max = 10,
       main = paste(capitalize(i),"of rules of products bought by retailers"))
  plot(rules_retcat, method = i,control=list(type="items"), max = 10,
       main = paste(capitalize(i),"of rules of categories bought by retailers"))
  
}
```